Is it time for an NBA expansion?

### 2024-04-20 paper/paper.qmd
Error Fixes

library(tidyverse)
library(knitr)
library(modelsummary)
library(tinytex)
library(lubridate)
```

@@

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| warning: false
#| message: false
#| label: fig-predict
#| fig-cap: PPG Predecition Through the 2027 Season

@@

create_future_data <- function(additional_teams) {
  latest_data <- tail(nba_data, 1)
  adjustment_factor <- 1 - 0.03 * additional_teams  # Each new team reduces stats by 3%
  data.frame(
    Year = c(2024, 2025, 2026),
    Num_Teams = latest_data$Num_Teams + additional_teams,
    AST = rep(latest_data$AST * adjustment_factor, 3),
    TRB = rep(latest_data$TRB * adjustment_factor, 3),
    STL = rep(latest_data$STL * adjustment_factor, 3),
    BLK = rep(latest_data$BLK * adjustment_factor, 3),
    TOV = rep(latest_data$TOV * adjustment_factor, 3)
  )
}

@@

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

Our results are summarized in @tbl-modelresults.

@@

line_colors <- c("red", "green", "blue", "purple", "orange", "cyan")

@@

combined_plot <- ggplot(nba_data_long, aes(x = Year, y = Value)) +
  geom_point(color = "black") +  # All points in black
  geom_smooth(method = "lm", aes(color = Statistic), se = FALSE) +  # Lines of best fit with different colors
  scale_color_manual(values = line_colors) +
  geom_vline(xintercept = exp_year, color = "black", linetype = "dashed") +
  facet_wrap(~Statistic, scales = "free_y") +  # Facet by statistic, with free y scales
  labs(title = "NBA Stats Over Years", x = "Year", y = "Value") +
  theme_minimal() +
  theme(legend.position = "none")

@@

@@

ggplot(scoring_data, aes(x = Year, fill = factor(Year))) +
  geom_histogram(stat = "count", binwidth = 1, color = "black") +
  labs(title = "Frequency of Records by Year",
       x = "Year",
       y = "Count of Records",
       fill = "Year") +
  theme_minimal() +
  theme(legend.position = "none")
##
data_1980_onwards <- scoring_data %>%
  filter(Year >= 1980)

# Create histogram for count of records per year from 1980 onwards
ggplot(data_1980_onwards, aes(x = Year)) +
  geom_histogram(stat = "count", bins = length(unique(data_1980_onwards$Year)), fill = "blue", color = "black") +
  labs(title = "Count of Records per Year from 1980 Onwards",
       x = "Year Modified",
       y = "Count of Records") +
  theme_minimal()

##
data_1980_60pts_plus <- scoring_data %>%
  filter(Year >= 1980, PTS >= 70)

# Create histogram for count of records per year for scores of 60 or more from 1980 onwards
ggplot(data_1980_60pts_plus, aes(x = Year)) +
  geom_histogram(stat = "count", bins = length(unique(data_1980_60pts_plus$Year)), fill = "red", color = "black") +
  labs(title = "Count of Records per Year for Scores of 70+ PTS from 1980 Onwards",
       x = "Year Modified",
       y = "Count of Records") +
  theme_minimal()
```

@@

msummary(
  list(
    "Points Model" = nba_model
  ),
  fmt = 2
  )
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| 
summary(nba_model)
```

>>>>

library(tidyverse)
library(knitr)
library(modelsummary)
library(tinytex)
library(lubridate)
library(viridis)
```

@@

@@

```{r}
#| warning: false
#| message: false
#| include: false
#| label: fig-predict
#| fig-cap: PPG Predecition Through the 2027 Season

@@

create_future_data <- function(additional_teams) {
  latest_data <- tail(nba_data, 1)
  adjustment_factor <- 1 - 0.03 * additional_teams  # Each new team reduces stats by 3%
  data.frame(
    Year = c(2025, 2026, 2027),
    Num_Teams = latest_data$Num_Teams + additional_teams,
    AST = rep(latest_data$AST * adjustment_factor, 3),
    TRB = rep(latest_data$TRB * adjustment_factor, 3),
    STL = rep(latest_data$STL * adjustment_factor, 3),
    BLK = rep(latest_data$BLK * adjustment_factor, 3),
    TOV = rep(latest_data$TOV * adjustment_factor, 3)
  )
}

@@

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-nbastats
#| fig-cap: NBA Statistics from 1980

@@

line_colors <- c("darkgreen", "mediumseagreen", "turquoise", "deepskyblue", "dodgerblue", "darkblue")

@@

combined_plot <- ggplot(nba_data_long, aes(x = Year, y = Value, group = Statistic)) +
  geom_point(color = "black") +  # All points in black
  geom_line(aes(color = Statistic)) +  # Connect each point with a line per statistic
  scale_color_manual(values = line_colors) +
  geom_vline(xintercept = exp_year, color = "black", linetype = "dashed") +
  facet_wrap(~Statistic, scales = "free_y") +  # Facet by statistic, with free y scales
  labs(title = "NBA Stats Over Years", x = "Year", y = "Value") +
  theme_minimal() +
  theme(legend.position = "bottom")  # Adjusted to show the legend

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-newnbastats
#| fig-cap: NBA Stats Since 2004

@@

ggplot(scoring_data, aes(x = Year, fill = factor(Year))) +
  geom_histogram(stat = "count", binwidth = 1, color = "black") +
  scale_fill_viridis(discrete = TRUE, option = "mako") +  # Using 'magma' for a warm color palette
  labs(title = "Frequency of Records by Year",
       x = "Year",
       y = "Count of Records",
       fill = "Year") +
  theme_minimal() +
  theme(legend.position = "none")

```{r}

data_1980_onwards <- scoring_data %>%
  filter(Year >= 1980)

ggplot(data_1980_onwards, aes(x = Year, fill = as.factor(Year))) +
  geom_histogram(stat = "count", bins = length(unique(data_1980_onwards$Year)), color = "black") +
  scale_fill_viridis(discrete = TRUE) +  # Apply viridis color scale for discrete data
  labs(title = "Count of Records per Year from 1980 Onwards",
       x = "Year Modified",
       y = "Count of Records") +
  theme_minimal() +
  theme(legend.position = "none")

```{r}
##
data_1980_70pts_plus <- scoring_data %>%
  filter(Year >= 1980, PTS >= 70)

ggplot(data_1980_70pts_plus, aes(x = Year, fill = factor(Year))) +
  geom_histogram(stat = "count", bins = length(unique(data_1980_70pts_plus$Year)), color = "black") +
  scale_fill_viridis_d(begin = 0.2, end = 0.8, direction = 1, option = "magma") +  # Discrete viridis scale
  labs(title = "Count of Records per Year for Scores of 70+ PTS from 1980 Onwards",
       x = "Year Modified",
       y = "Count of Records") +
  theme_minimal() +
  theme(legend.position = "none")
```

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig-cap: Linear NBA Model

msummary(
  list("Points Model" = nba_model),
  fmt = 2
  )
```
